---
title: "COVID-19 EDA"
author: "Rashmi Patel"
date: "6/14/2021"
output: html_document
---

```{r}
library(tidyverse)
library(DataExplorer)
library(covid19.analytics)
```
```{r}
covid19=covid19.JHU.data()
head(covid19)
dim(covid19)
unique(covid19$Country_Region)
str(covid19)
```
```{r}
# Checking the dimensions of the dataset
dim(covid19)
# Checking the column names of the dataset
colnames(covid19)
# Looking at the summary of the dataset
summary(covid19)
# Checking for data types of the columns of the dataset
str(covid19)
# Checking for number of columns with numeric type
numeric_var_who=sum(sapply(covid19[,1:14],is.numeric))
numeric_var_who
# Checking for number of columns with character type
char_var_who=sum(sapply(covid19[,1:14],is.character))
char_var_who
# Checking for column names with numeric type
numeric_varname_who=which(sapply(covid19[,1:14],is.numeric))
numeric_varname_who
# Checking for column names with character type
char_varname_who=which(sapply(covid19[,1:14],is.character))
char_varname_who

```
```{r}
plot_density(covid19)

sum(!is.na(covid19))
```

# Extracting Covid-19 data for United States of America from the total data
```{r}
usa.covid=covid19%>%filter(Country_Region=="US")
head(usa.covid,20)
tail(usa.covid,20)
```

# Covid-19 data for each state in United States of America
```{r}
state.total=usa.covid%>%group_by(Province_State)%>%summarise(usa.sum=sum(Confirmed))
sum(state.total$usa.sum)
ggplot(state.total,aes(x=reorder(as.factor(Province_State),-usa.sum),y=usa.sum))+geom_col(fill="sky blue",col="black")+
   theme(axis.text.x = element_text(angle = 90))+
   labs(x="State",y="Total cases")+ggtitle("Graph for total cases for each state in United States of America")

state.total%>%dplyr::select(Province_State,usa.sum)%>%arrange(desc(usa.sum))

```

# Covid-19 data for California by each County
```{r}
california=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="California")

ggplot(california,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+
   theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in California")

california%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Texas by each County
```{r}
texas=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Texas")

ggplot(texas,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+
   theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Texas")

texas%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Florida by each County
```{r}
florida=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Florida")

ggplot(florida,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Florida")

florida%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for New York by each County
```{r}
newyork=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="New York")

ggplot(newyork,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in New York")

newyork%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Illinois by each County
```{r}
illinois=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Illinois")

ggplot(illinois,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Illinois")

illinois%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Pennsylvania by each County
```{r}
pennsylvania=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Pennsylvania")

ggplot(pennsylvania,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Pennsylvania")

pennsylvania%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Georgia by each County
```{r}
georgia=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Georgia")

ggplot(georgia,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Georgia")

georgia%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Ohio by each County
```{r}
ohio=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Ohio")

ggplot(ohio,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Ohio")

ohio%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for New Jersey by each County
```{r}
newjersey=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="New Jersey")

ggplot(newjersey,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in New Jersey")

newjersey%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for North Carolina by each County
```{r}
northcarolina=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="North Carolina")

ggplot(northcarolina,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in North Carolina")

northcarolina%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for South Carolina by each County
```{r}
southcarolina=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="South Carolina")

ggplot(southcarolina,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in South Carolina")

southcarolina%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for New Mexico by each County
```{r}
newmexico=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="New Mexico")

ggplot(newmexico,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in New Mexico")

newmexico%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Alabama by each County
```{r}
alabama=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Alabama")

ggplot(alabama,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Alabama")

alabama%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Alaska by each County
```{r}
alaska=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Alaska")

ggplot(alaska,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Alaska")

alaska%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Arizona by each County
```{r}
arizona=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Arizona")

ggplot(arizona,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Arizona")

arizona%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Arkansas by each County
```{r}
arkansas=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Arkansas")

ggplot(arkansas,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Arkansas")

arkansas%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Colorado by each County
```{r}
colorado=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Colorado")

ggplot(colorado,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Colorado")

colorado%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Delaware by each County
```{r}
delaware=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Delaware")

ggplot(delaware,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Delaware")

delaware%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

usa.covid
```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```
# Covid-19 data for Connecticut by each County
```{r}
connecticut=usa.covid%>%group_by(Province_State)%>%filter(Province_State=="Connecticut")

ggplot(connecticut,aes(x=reorder(as.factor(Admin2),-Confirmed),y=Confirmed))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

connecticut%>%dplyr::select(Admin2,Confirmed)%>%arrange(desc(Confirmed))

```


# Covid-19 data deaths for US 
```{r}
usa.deaths=usa.covid%>%group_by(Province_State)%>%summarise(Deaths=sum(Deaths))
head(usa.deaths)
ggplot(usa.deaths,aes(x=reorder(as.factor(Province_State),-Deaths),y=Deaths))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

usa.deaths%>%dplyr::select(Province_State,Deaths)%>%arrange(desc(Deaths))


```
# Covid-19 data recovery for US 
```{r}
usa.recovery=usa.covid%>%group_by(Province_State)%>%summarise(Recovered=sum(Recovered))
head(usa.recovery)
ggplot(usa.recovery,aes(x=reorder(as.factor(Province_State),-Recovered),y=Recovered))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="County",y="Total cases")+ggtitle("Graph for total cases for each county in Connecticut")

usa.recovery%>%dplyr::select(Province_State,Recovered)%>%arrange(desc(Recovered))

sum(is.na(usa.covid$Recovered))
```
# Retreiving the USA population dataset
```{r}
library(usmap)
library(ggplot2)
library(dplyr)

state=read.csv("https://raw.githubusercontent.com/RashmiAPatel19/SMU-7330-DBMS/main/Term%20Project/csvData.csv",header=TRUE)
head(state)
dim(state)
colnames(state)
usa.data=usa.covid%>%group_by(Province_State)%>%summarise(Confirmed=sum(Confirmed),Deaths=sum(Deaths))
dim(usa.data)
df=subset(usa.data,Province_State!="Diamond Princess"&Province_State!="Grand Princess"&Province_State!="North Mariana Islands"&Province_State!="Puerto Rico"&Province_State!="Recovered"&Province_State!="Virgin Islands"&Province_State!="Guam")
dim(df)

population=data.frame(State=state$State,Population=state$Pop,ConfirmedCases=df$Confirmed,DeathsCases=df$Deaths)
head(population)


```
```{r}
ggplot(population,aes(x=reorder(as.factor(State),-Population),y=Population))+geom_col(fill="maroon",col="black")+theme(axis.text.x = element_text(angle = 90))+
   labs(x="State",y="Population")+ggtitle("Graph for population for each ")

population%>%arrange(desc(Population))


```
